Tabular Constraint Learning

### 2016-04-07 Paper/ECAI/tabular.tex
writing the formalization

\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $G$ -- set of groups
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
    \State $S \gets \emptyset$ \Comment{The set of solutions}
  \For{$c \in \constraints$} \Comment{\constraints -- the set of predefined Excel constraints}
  \State $v_1, ..., v_n =$ variables of $c$ 
    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S)$}
      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S)$
    \EndFor
  \EndFor\\
\Return $S$
\end{algorithmic}
\caption{Tabular constraint learning}
\label{algo:tcl}
\end{algorithm}

\sergey{we assume an order on the constraints (it doesn't have to be complete though)}

\subsection{Groups}
In this work we distinguish the following types of data: numeric and textual. The numeric type has two subtypes: integers and floats. We also consider the special element called \textit{None}, which has two types: numeric and textual. A set is called type-consistent iff all elements are either numeric or textual. Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.

@@

\subsection{Algorithm description}

Let us describe the key functions in Algorithm \ref{algo:tcl}.
\paragraph{$\generategroups(\textit{Constraint,GroupSet,Solutions})$} is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = SUM(X)} and $X,Y$ satisfy the following constraints \sergey{Samuel, can you add here, what is actually used for the sum now?}. Essentially, the group generation step is a constraint satisfaction problem associated with the specific constraint. However, many constraints have the same candidate generation procedures, e.g., min, max, avg, count, etc.

\paragraph{$\findassignment(\textit{Constraint,Candidates,Solutions})$} is the function looking for the subsets of vectors in the candidates satisfying the constraint. If multiple subsets satisfy the constraint, a maximal is selected. If $G_1,G_2$, associated with the variables $X,Y$ in \textit{Y = SUM(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and consecutive subset of vectors $x$ in $G_1$ such that the sum over $x$ is equal to $y$. For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}7]$ can be used for both $X$ and $Y$ in the row-wise sum constraint, then as $x$ would be selected $T_1[{:},3{:}6]$ and as $y$ would be $T_1[{:},7]$.

\paragraph{}

>>>>

\newcommand{\CName}{Name\xspace}
\newcommand{\CSignature}{Signature\xspace}
\newcommand{\CFunction}{Function\xspace}

\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $G$ -- set of groups
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
    \State $S \gets \emptyset$ \Comment{The set of solutions}
  \For{$c \in \constraints$} \Comment{\constraints -- partially ordered set of Excel constraints; Table \ref{table:constraints}}
  \State $v_1, ..., v_n =$ variables of $c$ 
    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S)$}
      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S)$
    \EndFor
  \EndFor\\
\Return $S$
\end{algorithmic}
\caption{Tabular constraint learning}
\label{algo:tcl}
\end{algorithm}
\subsection{Groups, Type-consistency and Constraints}
In this work we distinguish the following types of data: numeric and textual. The numeric type has two subtypes: integers and floats. We also consider the special element called \textit{None}, which has two types: numeric and textual. A set is called \textit{type-consistent} iff all elements are either numeric or textual. Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.

@@

A \textit{constraint} is a triple of \textit{(\CName,\CSignature,\CFunction)}, where \CName is a textual name of the constraint; \CSignature is the signature of the constraint consisting of its arity and properties of the arguments, such as their types, sizes (properties might constrain sizes of two arguments as well, by for example, require them to be equal), etc; \CFunction is the function over arguments specified in \CSignature that represents constraint satisfaction. For example, a column-wise sum constraint has \CName equal to \textit{Y = SUM(X, col)}, its \CFunction sets arity to 2 and both arguments must be numeric and \sergey{Samuel, could you put the actual properties and \CFunction here?}

\subsection{Constraint learning algorithm description}
Let us describe the key steps and functions in Algorithm \ref{algo:tcl}. Essentially, the algorithm has two steps: candidate group generation and subgroup satisfaction search, which is in line with the well-known in AI paradigm of ``generate-and-test'' \cite{whaisasp}. Let us elaborate on each step in detail.

\paragraph{Candidate group generation} $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = SUM(X)} and $X,Y$ satisfy the following constraints \sergey{Samuel, can you add here, what is actually used for the sum now?}. Essentially, the group generation step is a constraint satisfaction problem associated with the specific constraint. However, many constraints have the same candidate generation procedures, e.g., min, max, avg, count, etc.

\paragraph{Subgroup satisfaction search} $\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function looking for the subsets of vectors in the candidates satisfying the constraint. If multiple subsets satisfy the constraint, a maximal is selected. If $G_1,G_2$, associated with the variables $X,Y$ in the constraint \textit{Y = SUM(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and consecutive subset of vectors $x$ in $G_1$ such that the sum over $x$ is equal to $y$. For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}7]$ can be used for both $X$ and $Y$ in the row-wise sum constraint, then as $x$ would be selected $T_1[{:},3{:}6]$ and as $y$ would be $T_1[{:},7]$.

\subsection{Constraints}
\sergey{we need to explain that constraints order is a partial order, which is a DAG}
\begin{table}
  \centering
  \begin{tabular}{lcc}
    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ 
  \end{tabular}
  \caption{Tabular Constraints}
  \label{table:constraints}
\end{table}
